---
title: "Onet"
author: "simplymathematics"
date: "October 19, 2018"
output: html_document
---

# Dependencies

```{r}
require(XML) #for reading xml, html
require(stringr) #for regex
require(curl) #for downloading file
require(rvest) #for extracting text from a webpage
```

# Downloading
We used the Onet data to cross references our findings from before. In particular, we looked at the [bright outlook occpations](https://www.onetonline.org/find/bright?b=0&g=Go) that had a math prefix. There were six, stored here are -urls.- We also saved the files as onet[1-6].html. 

```{r}
urls <- c(
  "https://www.onetonline.org/link/summary/15-2011.00",
  "https://www.onetonline.org/link/summary/15-2041.01",
  "https://www.onetonline.org/link/summary/15-2041.02",
  "https://www.onetonline.org/link/summary/15-1111.00",
  "https://www.onetonline.org/link/summary/15-1151.00",
  "https://www.onetonline.org/link/summary/15-1141.00"
)
i <- 1 #initialize a counter
for (url in urls){
  file = paste(c("onet",i,".html"),collapse = "")
  i <- i + 1
  curl_download(url, file)
}

numbers.list <- c(2011, 2041, 1111, 1151, 1141)
```
Below, we're importing each of these files into a dataframe with a human friendly name. The final vector, -jobs,-are the occupations we'll focus on.
```{r}
actuary <- htmlTreeParse("onet1.html")
biostatitician <- readHTMLTable("onet2.html")
clinical.data.manager <- readHTMLTable("onet3.html")
computer.research.scientist <- readHTMLTable("onet4.html")
user.support.specialist <- readHTMLTable("onet5.html")
database.administrator <- readHTMLTable("onet6.html")
jobs <- c(actuary, biostatitician, clinical.data.manager, computer.research.scientist, user.support.specialist, database.administrator)
```
For now, we'll start with one job.
```{r}
actuary
```

I am trying to pull the skills data for each occupation. For some reason, I can not get R to recognize "Skills" or "Technical Skills". You can see in the summary below that the readHTMLtable and the htmltreeparse are only picking out 4 of the many tables listed on the website. I will keep on this to find a way to pull out the neccessary data. - jm

```{r}
typeof(actuary)
length(actuary)
str(actuary)
```
