---
title: "DB Creation SQLite fm CSV-Version2"
output: html_document
---


```{r}
library(RSQLite)
  # need to add this library first
library(sqldf)
library(dplyr)
library (tidyverse)
```


```{r}
#setwd ('E:/Proj3 Nov3 Charlie FolderGithub/csvs')
setwd ('C:/Users/User/Desktop/To Email DeskTop')
#E:\Proj3 Nov3 Charlie FolderGithub\csvs
# Physical library on Gracie's PC
getwd()

#try(setwd("csvs"))
# Relative path in Github
```

 Database1: Start with Indeed Data Scraping Data
 
```{r}
# Load csv file containing foreign key into a dataframe
Keyword_Search_csv_df <- read.csv("Indeed_Keyword_Search_data.csv")

# Load csv file containing primary key into a dataframe
keywords_csv_df <- read.csv("Indeed_Keyword.csv")

# Select relevent columns for database table insert
Keyword_Search_csv_df <- Keyword_Search_csv_df %>% 
  select(Data_Scientist_Skills_ID, Company, Web_scrape_date, state, job_post_link)
head(Keyword_Search_csv_df)

# Select relevent columns for database table insert
keywords_csv_df <- keywords_csv_df %>% 
  select(ID, Keyword)
head(keywords_csv_df)
```

Database2: continue with the ONET (Labor Dept Data)
ONET Characteristics table is the main table from ONET database, then Salary table, then Skills (Soft) table
```{r}
characteristics <- read.csv ("characteristics.csv")

characteristics <- characteristics %>% 
  select (Title,SOC ,	OccupationType, 	TypicalEntryLvlEduc,	
           PreEmplExperience,	PostEmplTraining)
head(characteristics)


salary <- read.csv ("salary.csv")
salary <- salary %>% 
  select (	SOC,	No.Employees,	RSE,	Mean.Hourly.Wage,	Mean.Annual.Wage,	Wage.RSE)
head(salary)
```



The linkage bw ONET and INDEED data is the technical skills 
## Tech Skill format 1: wide format, each skill is spreaded
```{r}

skills_sum <- read.csv ('skills_sum.csv')
skills_sum %>% 
  select (jobNoNA,	jobcodeNoNA,		C,		Django,         	JavaScript,
          MongoDB,	MySQL,		NoSQL,         	PostgreSQL,	Python  ,
          R,	Ruby          ,SAS,	         	STATISTICA,	          SuperANOVA,	Tableau	)
#head(skills_sum)
str(skills_sum)

```

## Tech Skill format 2: Long format, each skill is gathered
```{r}
TechSkillswOccupation <- read.csv("skills_sumcob_TechSkillswOccupation.csv")
TechSkillswOccupation %>% 
  select (Section,	Category,	Example,	job	,jobcode)
#head (TechSkillswOccupation)
str(TechSkillswOccupation)
```

Connect to the database
## we create a temporary SQLite database called Proj3_67 (changed name to mydb)
```{r}
mydb <- dbConnect(RSQLite::SQLite(), "Proj3_607.db")
# List out how many tables are in this object
 dbListTables(mydb)
 
# TEST, to get query from this database (ONET Characteristic table only for now)
dbGetQuery  (mydb,"Select* FROM C") 
```

Relational Database 1st set
#  Two relational Tables within Indeed Web Scraping, linked by Primary Key and Foreigh Key
```{r}

# Create/Populate tables with data from the dataframes
dbWriteTable(mydb, "webscrapes", Keyword_Search_csv_df, overwrite = TRUE)
dbWriteTable(mydb, "keywords", keywords_csv_df, overwrite = TRUE)

# Join the two tables using the primary and foreing key(relational database)
join_tb_indeed <- "select keywords.Keyword, webscrapes.Company, webscrapes.Web_scrape_date,webscrapes.state 
from webscrapes 
inner join keywords
on keywords.ID = webscrapes.Data_Scientist_Skills_ID"

head(dbGetQuery(mydb, join_tb_indeed))
tail(dbGetQuery(mydb, join_tb_indeed))

```

 Relational Database 2nd set:  
#Two relational Tables within ONET (Labor dept) Web Scraping, linked by Primary Key and Foreigh Key
```{r}

# ONET Characteristics table is the main table from ONET database,
dbWriteTable(mydb,"C", characteristics, overwrite = TRUE)  
dbWriteTable(mydb,"salary", salary, overwrite = TRUE)

join_tb_onetsalary <- "select C.title, C.SOC, 	C.OccupationType, 	C.TypicalEntryLvlEduc,	
           C.PreEmplExperience,	C.PostEmplTraining,
            salary.SOC,  RSE
from C
inner join salary
on C.SOC = salary.SOC"

head (dbGetQuery(mydb,join_tb_onetsalary))

```

Join the ONET SKILLS (soft Skills) with ONET Characteristics TABLES

```{r}

dbWriteTable(mydb,"C", characteristics, overwrite = TRUE)
dbWriteTable(mydb, "TechSkl", TechSkillswOccupation, overwrite = TRUE)
join_tb_onetTechSkl <- 
  "select C.title, C.SOC, 	C.OccupationType, 	C.TypicalEntryLvlEduc,	
           C.PreEmplExperience,	C.PostEmplTraining,
            TechSkl.job, TechSkl.jobcode,TechSkl.Example
from C
inner join  TechSkl
on C.Title = TechSkl.job"

head (dbGetQuery(mydb,join_tb_onetTechSkl))

```

 Relational Database 3rd set: Link INDEED web scraping and Labor Dept ONET webscraping
```{r}
# Linkage between those two sets: using the Technical Skills 
# Tech Skills linked with ONET Database
join_tb_ONETTechSkl <- 
  "select C.title, C.SOC, 	C.OccupationType, 	C.TypicalEntryLvlEduc,	
           C.PreEmplExperience,	C.PostEmplTraining,
            TechSkl.job, TechSkl.jobcode,TechSkl.Example
from C
inner join  TechSkl
on C.Title = TechSkl.job"
head (dbGetQuery(mydb,join_tb_ONETTechSkl))

```

Tech Skills linked with Indeed Database
```{r}
join_tb_IndeedTechSkl <- "select keywords.Keyword, 
 TechSkl.job, TechSkl.jobcode,TechSkl.Example
from TechSkl
inner join keywords
on keywords.Keyword = TechSkl.Example"
head (dbGetQuery(mydb,join_tb_IndeedTechSkl))

```
List all the relational tables within the SQlite database
```{r}
dbListTables(mydb)
# all tables are displayed here, 6 tables (all relational in total)
```

List the column fields within each individual tables within the SQLite Database
```{r}
dbListFields(mydb,"C")
```

```{r}
dbListFields(mydb,"SftSkl")
```

```{r}
dbListFields(mydb,"TechSkillswOccupation")
```

```{r}
dbListFields(mydb,"salary")

```

```{r}
dbListFields(mydb,"TechSkl")

```

```{r}
dbListFields(mydb,"keywords")

```

```{r}
dbListFields(mydb,"webscrapes")
```

```{r}
setwd('..')
```

