---
title: "DB Creation SQLite fm CSV-Version2"
output: html_document
---

# Dependencies

```{r}
library (tidyverse)
library(RSQLite)
library(DBI)
```

```{r}
try(setwd(dir = "csvs/"))
```

# Connect to the database
```{r}
data.skills.db <- dbConnect(SQLite(), "db.sqlite")
dbListTables(data.skills.db)
```

# Load Data: Here, we load the indeed data into indeed dataframes.
 
```{r}
# Load csv file containing foreign key into a dataframe
indeed.ca <- read.csv("indeed_keyword_count_ca.csv")
indeed.ny <- read.csv("indeed_keyword_count_ny.csv")
```

# Here, we load the variosu ONET/BLS data into various objects.
```{r}
characteristics <- read.csv ("characteristics.csv")
salary <- read.csv ("salary.csv")
skills <- read.csv ('skills.csv')
tech.skills <- read.csv('TechSkills.csv')
```

# Here we write the Indeed data as two se
```{r}
# Create/Populate tables with data from the dataframes
dbWriteTable(data.skills.db,"indeed.ca", indeed.ca, overwrite = TRUE)
dbWriteTable(data.skills.db,"indeed.ny", indeed.ny, overwrite = TRUE)
```
# Write Data

```{r}
dbWriteTable(data.skills.db,"characteristics", characteristics, overwrite = TRUE)  
dbWriteTable(data.skills.db,"salary", salary, overwrite = TRUE)
dbWriteTable(data.skills.db,"skills", skills, overwrite = TRUE)
dbWriteTable(data.skills.db,"tech.skills", tech.skills, overwrite = TRUE)
```

# Check in on the database
 
```{r}
dbListTables(data.skills.db)
```

The SQLite documentation says that, 

> It is not possible to use the "ALTER TABLE ... ADD COLUMN" syntax to add a column that includes a REFERENCES clause, unless the default value of the new column is NULL. Attempting to do so returns an error. 

This means that we would need to relate these through another means. Given time restrictions and the fact that our related data is all binned by a 4 digit SOC code, it would be trivial to import into MySQL Workbench and add the foreign keys.

